Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11- 10150.}

### 2018-06-25 2018/Journal of Supercomputing/JoS_Paper.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

@@

\title{Parallel Computing for Constrained Global Optimization Problems
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\Russian
Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными. Например, моделируемым объектом является некоторое электронное устройство, и минимизируется время его срабатывания. Однако время срабатывания является неопределенным, если при данном сочетании параметров устройство вообще не работает.

%Another example is the optimal control problem described through system of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the criteria is indeterminate.

@@

\Russian
Сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы - приведено в статье \cite{Barkalov2017}. В ней экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции. Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области). Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. 
%For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

%Проверить
For example, a scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}. The building of a Peano curves set by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme. In the initial non-rotated evolvent for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations might appear to be redundant. In this case, we can select only a part of all rotations. As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes. Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
\Russian
Именно такой подход к построению set of evolvents был использован в данной статье при проведении вычислительных экспериментов.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. 
\Russian
Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области. При этом можно задавать объем допустимой области и  число ограничений, активных в точке минимума, тем самым усложняя или упрощая постановку задачи.

@@

The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$. For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).

@@

\Russian
В заключение отметим, что подавляющее большинство оптимизационных алгоритмов (за исключением методов, основанных на идеях случайного поиска) реализуют стратегию целенаправленного выбора. Указанная стратегия состоит в том, чтобы на основе анализа небольшой части вариантов исключить из дальнейшего рассмотрения многие заведомо не перспективные случаи и сосредоточить дальнейший поиск в подмножестве, содержащем лучший вариант.

Сложность моделируемых процессов и объектов может приводить к тому, что их характеристики не будут обладать свойством монотонности, что обуславливает недостаточность методов локального поиска для оценки наилучшего варианта. В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска, например, за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров. Математическая формулировка этого факта может иметь форму Lipschitz condition, uniform H\"older condition и т.п. Именно к таким методам, способным решать сложные прикладные задачи оптимизации, относится рассмотренный в данной статье параллельный индексный алгоритм.

The parallel index method for solving constrained global optimization problems %considered in this paper 
offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed. This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.

\Russian
Перспективным направлением дальнейшей работы будет являться применение set of evolvents для редукции размерности в задачах многокритериальной оптимизации. В подобных задачах построение множества Парето требует решения целой серии скалярных задач. При их параллельном решении можно ожидать более высокого ускорения, чем при решении одной скалярной задачи.

@@

\bibitem{Korosec}
Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--772 (2013)

@@

%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex 
%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

%\usepackage[utf8x]{inputenc}
%\usepackage[english,russian]{babel}

@@

\title{Parallel Computing for Constrained Global Optimization Problems
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

This situation is typical for many optimal design problems, when at violation of some conditions 
of functioning of the modeled object, other characteristics of this one appear to be undefined. 
For example, the modeled object is an electronic device, and its response time is minimized. 
However, the response time is undefined if at given combination of parameters the device is not 
working at all.

%Another example is the optimal control problem described through system of ordinary 
%differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only 
%possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz 
%matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the 
%criteria is indeterminate.

@@

The comparison of two different methods of accounting for the constraints -- the classical 
penalty functions method and the index scheme -- is presented in \cite{Barkalov2017}. In 
this paper, the advantage of the index method over the penalty functions one in the number of 
iterations as well as in the number of computations of the constraints and of the objective 
function has been confirmed experimentally. To solve a problem with given accuracy, the index 
method required 3-5 times less iterations (subject to the problem dimensionality and the feasible 
domain fraction). At the same time, the objective function value was computed in average at 
every fifth iteration of the index method only, in difference from the penalty functions method, 
in which the objective function values were computed at every iteration.

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping 
$Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set 
$Y^S(x)$, different approaches can be used. 
%For example, in \cite{Strongin2000} a scheme was implemented, according to which each 
%evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent 
%$y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed 
%allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 
%1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in 
%one coordinate.

For example, a scheme of building of the multiple mapping was proposed in \cite{6_Gergel2009}. The building of a set of Peano curves by rotation of the evolvents 
around the coordinate origin is a distinctive feature of this scheme. In the initial non-rotated 
evolvent for close points $y', y''$ in the multidimensional space their preimages $x', x''$ in the 
interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping 
$y^i(x)$ according to which the preimages $x', x''$ will be located closer to each other. The 
maximum number of various rotations of the evolvents mapping the $N$-dimensional 
hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations 
might appear to be redundant. In this case, we can select only a part of all rotations. As a 
possible approach, we propose to generate the new evolvents by the rotation of the initial 
evolvent on the angle of $\pm\pi/2$ in each coordinate plane. Taking into account the initial 
evolvent, the total number of such evolvents for mapping the $N$-dimensional domain onto the 
corresponding one-dimensional intervals equals to $N(N-1)+1$. This method for building a set 
of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
Just such an approach to building the set of evolvents was employed in the present paper when 
carrying out the numerical experiments.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality 
proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating 
constrained global optimization problems is proposed in \cite{Gergel2017}. 
The test problems with constraints obtained using this scheme have an important property 
inherent to the applied optimization problems: the constrained global minimum is located at the 
feasible domain boundary whereas the global minimum of the objective function without 
constraints is located outside the feasible domain. At that, one can specify the feasible domain 
volume and the number of constraints active at the minimum point thus complicating or 
simplifying the problem statement.

@@

The experiments have been carried out by solving a series of $100$ problems with two 
constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS 
classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ 
and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at 
each node have been varied. The problem was considered to be solved, if the algorithm 
generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-
y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$. For the purpose of simulation of the computational 
complexity inherent to applied optimization problems, calculation of the problem functions in all 
performed experiments was made more complex by additional calculations without changing 
the type of function and arrangement of its minima (a series summation of $100$ thousand 
elements).

@@

In conclusion, note that the majority of the optimization algorithms (except the methods based 
on the random search ideas) implement the purposeful search strategy. This strategy consists in 
excluding many non-prospective cases from further consideration on the base of 
analysis of a small part of variants and concentrating further search in the subset containing the 
best variant.

The complexity of the modeled processes and objects can result in the non-monotonous
characteristics of these ones that causes the inadequacy of local search 
methods for evaluating the optimal variant. In such problems, the global search methods should 
be applied, which provide the purposeful search, for example, by means of accounting for the 
limited variation of the characteristics of the modeled object at limited variations of its 
parameters. The mathematical formulation of this fact can take the forms of the Lipschitz 
condition, uniform H\"older condition, etc. The parallel index algorithm considered in the 
present paper belongs just to such methods capable to solve the complex applied optimization 
problems.

The parallel index algorithm %for solving constrained global optimization problems %considered in this paper 
offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus 
eliminating the need to select the penalty coefficient and to solve a series of unconstrained 
problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined 
everywhere (for example, the objective function values are undefined outside the feasible 
domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory 
proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated a speedup with respect to the number of 
processors/cores employed. This was confirmed by the results of solving numerically of several 
hundred test problems using 160 cores of the Lobachevsky supercomputer.

The application of the set of evolvents for the dimensionality reduction in the multicriterial 
optimization problems would be a promising direction of further work. In such problems, the 
construction of Pareto set requires solving even a series of scalar problems. When solving these 
ones in parallel, one can expect a higher speedup than when solving a single scalar problem.

@@

\bibitem{Korosec}
Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the 
differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--
772 (2013)

@@

%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-
control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 
(2017)
